---
title: "Aids Drug"
author: "Tito Lopez"
date: "Fall 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aids drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

```{r}
library(dplyr)
druggies = filter(triUmp, triUmp$DrugUser== "yes")
```

$H_0$ There is no difference in how effective the drug works for aids

$H_A$ There is a difference in how effective the drug works for aids

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

The response variable is categorical. The response variable is the infected which is based on the treatment of the new aid drug. The explanatory variable is treatment which is categorical. Therefore we use cat~cat.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=druggies)
barchartGC(~infected + treatment,data=druggies, type="percent")
```

This bar graph shows people who has aids and took the drug did better than the people who didn't take the drug. The people who don't have aids who took the drug stayed fairly the same to those who didn't take the drug.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=druggies)
rowPerc(table1)
colPerc(table1)
```

This table shows the people who has aids that took the drug, 24% still had aids and people who took the placebo, 76% still had aids. The people without aids who took the drug, 48% got aids and the people who took the placebo, 52% got aids.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

```{r}
#druggies$infected = factor(druggies$infected, c('none','aids'))
druggies$treatment = factor(druggies$treatment,c('placebo','drug'))
table1 <-xtabs(~infected + treatment, data=druggies)
rowPerc(table1)
colPerc(table1)
```

The p value is 0.0000001242. Which is smaller than 0.05, therefore I will reject the null hypothesis. The people who have aids and take the drug are more likely to be cured than the people who take the placebo. The people who don't have aids who take the drug will more likely get aids. The people who don't have aids that take the placebo will have around the same results as the people who take the drug.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The question is Consider only the DrugUsers, does this drug seem effective? Yes this drug is effective because the people with aids who take the drug are more likely to not have aids anymore than the people who take the placebo. The p value is 0.0000001242. Since it is smaller than 0.05, I will reject null hypothesis. Therefore The people who have aids and take the drug are more likely to be cured than the people who take the placebo. The people who don't have aids who take the drug will more likely get aids. The people who don't have aids that take the placebo will have around the same results as the people who take the drug.
